---
title: "Iteration and listcols"
output: github_document
---

```{r setup, include = FALSE}
library(tidyverse)

library(rvest)
library(p8105.datasets)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.color = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_color_discrete = scale_color_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```


## Lists

You can put anything in a list

```{r}
vec_numeric = 5:8
vec_char = c("My", "name", "is", "Jeff")
vec_logical = c(TRUE, TRUE, TRUE, FALSE)
```

```{r}
l = list(
  vec_numeric = 5:8,
  mat         = matrix(1:8, 2, 4),
  vec_logical = c(TRUE, FALSE),
  summary     = summary(rnorm(1000)))
l
```

```{r}
l$vec_numeric
l[[1]]
l[[1]][1:3]
```

## `for` loop

Create a new list.

```{r}
list_norms = 
  list(
    a = rnorm(20, 3, 1),
    b = rnorm(20, 0, 5),
    c = rnorm(20, 10, .2),
    d = rnorm(20, -3, 1)
  )

is.list(list_norms)
```

Pause and get my old function.

```{r}
mean_and_sd = function(x) {
  
  if (!is.numeric(x)) {
    stop("Argument x should be numeric")
  } else if (length(x) == 1) {
    stop("Cannot be computed for length 1 vectors")
  }
  
  mean_x = mean(x)
  sd_x = sd(x)

  tibble(
    mean = mean_x, 
    sd = sd_x
  )
}
```

I can apply that function to each list element.

```{r}
mean_and_sd(list_norms[[1]])
```

```{r}
mean_and_sd(list_norms[[2]])
```

```{r}
mean_and_sd(list_norms[[3]])
```

```{r}
mean_and_sd(list_norms[[4]])
```

Let's use a for loop:

```{r}
output = vector("list", length = 4)

for (i in 1:4) {
  output[[i]] = mean_and_sd(list_norms[[i]])
}
```

## Let's try map!

```{r}
output = map(list_norms, mean_and_sd)
```

what if you want a different function...?

```{r}
output = map_dbl(list_norms, median, .id = "input")
```

```{r}
output = map_dfr(list_norms, mean_and_sd, .id = "input")
```

```{r, error = TRUE }
output = map2(input_1, input_2, \(x,y) func(arg_1 = x, arg_2 = y))
```


## List columns

```{r}
listcol_df = 
  tibble(
    name = c("a", "b", "c", "d"),
    samp = list_norms
  )
```

```{r}
listcol_df |> pull(name)
```

```{r}
listcol_df |> pull(samp)
```

```{r}
listcol_df$samp[[1]]
```

Lets' try some operations

```{r}
mean_and_sd(listcol_df$samp[[1]])
```

Can I just ... map?

```{r}
map(listcol_df$samp, mean_and_sd)
```


So ... can I add a list column?

```{r}
listcol_df = 
  listcol_df |> 
  mutate(summary = map(samp, mean_and_sd))

listcol_df
```


## Revisiting NSDUH

```{r}
nsduh_table <- function(html, table_num) {
  
  table = 
    html |> 
    html_table() |> 
    nth(table_num) |>
    slice(-1) |> 
    select(-contains("P Value")) |>
    pivot_longer(
      -State,
      names_to = "age_year", 
      values_to = "percent") |>
    separate(age_year, into = c("age", "year"), sep = "\\(") |>
    mutate(
      year = str_replace(year, "\\)", ""),
      percent = str_replace(percent, "[a-c]$", ""),
      percent = as.numeric(percent)) |>
    filter(!(State %in% c("Total U.S.", "Northeast", "Midwest", "South", "West")))
  
  table
}
```

```{r}
nsduh_url = "http://samhda.s3-us-gov-west-1.amazonaws.com/s3fs-public/field-uploads/2k15StateFiles/NSDUHsaeShortTermCHG2015.htm"

nsduh_html = read_html(nsduh_url)

output = vector("list", 3)

for (i in c(1, 4, 5)) {
  output[[i]] = nsduh_table(nsduh_html, i)
}

nsduh_results = bind_rows(output)
```

```{r}
nsduh_results = 
  map(c(1, 4, 5), nsduh_table, html = nsduh_html) |> 
  bind_rows()
```

```{r}
nsduh_results= 
  tibble(
    name = c("marj", "cocaine", "heroine"),
    number = c(1, 4, 5)) |> 
  mutate(table = map(number, \(num) nsduh_table(html = nsduh_html, num))) |> 
  unnest(cols = "table")

```


## Weather Data (Operations on nested data)

```{r}
library(p8105.datasets)
data("weather_df")
```

Get our list columns ..
```{r}
weather_nest = 
  nest(weather_df, data = date:tmin)

weather_nest
```

```{r}
weather_nest |> pull(name)
```

```{r}
weather_nest |> pull(data)
```

```{r}
unnest(weather_nest, cols = data)
```


Suppose I want to regress `tmax` on `tmin` for each station.

Let's write a function

```{r}
weather_lm = function(df) {
  lm(tmax ~ tmin, data = df)
}
```

```{r}
weather_lm(weather_nest$data[[1]])
```

What about a map ... ?

```{r}
map(weather_nest$data, weather_lm)
```

What about a map in a list column 

```{r}
map(weather_nest$data, \(df) lm(tmax ~ tmin, data = df))
```

```{r}
weather_nest = 
  weather_nest |> 
  mutate(models = map(data, weather_lm))

weather_nest
```


## Using iteration for data import

```{r}
full_df = 
  tibble(
    files = list.files("data/exp_data/"),
    path = str_c("data/exp_data/", files)
  ) %>% 
  mutate(data = map(path, read_csv)) %>% 
  unnest()
```

```{r}
tidy_df = 
  full_df %>% 
  mutate(
    files = str_replace(files, ".csv", ""),
    group = str_sub(files, 1, 3)) %>% 
  pivot_longer(
    week_1:week_8,
    names_to = "week",
    values_to = "outcome",
    names_prefix = "week_") %>% 
  mutate(week = as.numeric(week)) %>% 
  select(group, subj = files, week, outcome)
```

```{r}
tidy_df %>% 
  ggplot(aes(x = week, y = outcome, group = subj, color = group)) + 
  geom_point() + 
  geom_path() + 
  facet_grid(~group)
```



